### 2014-06-20 ms/ms.Rmd
add reference

# *Ecological Archives* Exxxxx
# Falster D.S. *et al* BAAD: a Biomass And Allometry Database for woody plants. *Ecology* xxxx

@@

library(plyr)
library(bibtex)
library(knitcitations)
library(Hmisc)
library(whisker)

@@

```{r,results="asis",echo=FALSE, message=FALSE}
writeLines(paste("**List of Authors:**", lisOfAuthors(baad$contact)))
```

@@

Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))`. The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`. While many studies in have destructively harvested woody plants in the name of science, most of these data have only been made available in the form of summary tables or figures included in publications. Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world. Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4","10.3832/ifor0901-006",bib["zianis_biomass_2005"]))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale. Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a"))`.

@@

Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518"))`. This compilation would not have been possible without several new technologies. In particular,

@@

@@

allometry; trait; leaf area; height; size; architecture; vegetation
model; carbon; growth; wood

@@

Count <- function(x)length(data[[x]][!is.na(data[[x]])])
N <- sapply(thesevars, Count)

dfr <- data.frame(Variable=Labels, N=N)
dfr <- dfr[dfr$N > 0,]
rownames(dfr) <- NULL
```

```{r results='asis', echo=FALSE}
kable(dfr, "html", align="l")
```

>>>>

---
output: html_document
---
# *Ecological Archives* Exxxxx
# Falster D.S. *et al* BAAD: a Biomass And Allometry Database for woody plants. *Ecology* xxxx

@@

library(plyr)
library(bibtex)
library(knitcitations)
library(Hmisc)
library(whisker)
library(scales)
library(maps)

@@

```{r,results="asis",echo=FALSE, message=FALSE}
writeLines(paste("**List of Authors:**", pasteC(listOfAuthors(baad$contact), ', ')))
```

@@

Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))`. The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`. While many studies in have destructively harvested woody plants in the name of science, most of these data have only been made available in the form of summary tables or figures included in publications. Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world. Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale. Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.

@@

Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518","10.1016/j.tree.2011.11.016"))`. This compilation would not have been possible without several new technologies. In particular,

@@

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=6}
drawWorldPlot  <-  function(data, horlines=TRUE, sizebyn=FALSE, add=FALSE,
                            pchcol="red",
                            legend=TRUE){

  if(!add){
    map('world',col="grey80",bg="white",lwd=0.5,fill=TRUE,resolution=0,wrap=TRUE, border="grey80")
    map('world',col="black",boundary=TRUE,lwd=0.2,interior=FALSE,fill=FALSE,add=TRUE,resolution=0,wrap=TRUE)
  }

  if(horlines){
    lines(c(-180,180),c(-100,-100), lty="dashed", xpd=TRUE, lwd=2)
    lines(c(-180,180),c(100,100), lty="dashed", xpd=TRUE, lwd=2)
  }

  # Remove all duplicates (increases speed and minimizes file size)
  latlon <- with(data, paste(latitude, longitude))
  lat <- data$latitude[!duplicated(latlon)]
  lon <- data$longitude[!duplicated(latlon)]

  j  <-  !is.na(lat) & !is.na(lon)

  # Location only sometimes missing - but lat/lon can still be in dataset anyway.
  # & data$loc != "NA" | is.na(data$loc)

  if(!any(j)){
    polygon(c(-100,95,95,-100), c(-10,-10,15,15), col=rgb(0,0,0,240,maxColorValue=255))
    text(-100, 0, expression(paste(bold("Missing coordinate/location"))), col="red", xpd=TRUE, pos=4, cex=0.8)
  } else {

    if(!sizebyn){
      points(lon,lat, pch=19, col=alpha("blue",0.5), cex=1.6)
    } else {
      n <- table(latlon)
      symbols(lon,lat, circles=log10(n), inches=0.1, fg="black", bg=pchcol, add=TRUE)

      if(legend){
        ns <- c(10,100,1000)
        X <- rep(-170,3)
        Y <- seq(-30,-10,by=10)
        rect(xleft=-200, xright=-120, ybottom=-50, ytop=10, col="white", border=NA)
        symbols(x=X, y=Y, circles=log10(ns), inches=0.1,
                fg="black", bg=pchcol, add=TRUE)
        text(x=X+5, y=Y, labels=as.character(ns),pos=4)
        text(x=X+5, y=Y[3]+10, labels=expression(italic(n)), pos=4)
      }

    }

  }

}

# Colored by vegetation type
  cols <- alpha(c("blue","brown","forestgreen","darkorange","dodgerblue2",
                    "navajowhite4","lightgoldenrod3","lavender"),0.67)
  vegs <- c("BorF","Sav","TempF","TempRF","TropRF","TropSF","Wo")
  vegs_labels <- c("Boreal forest","Savannah","Temperate forest",
                   "Temperate rainforest","Tropical rainforest",
                   "Tropical seasonal forest","Woodland","Glasshouse / Common garden")

  drawWorldPlot(data[data$vegetation == vegs[1],], horlines=FALSE,
                sizebyn=TRUE, pchcol=cols[1])
  for(i in 2:length(vegs)){
    drawWorldPlot(data[data$vegetation == vegs[i],], horlines=FALSE,
                  sizebyn=TRUE, pchcol=cols[i], add=TRUE)
  }
  drawWorldPlot(data[data$growingCondition == "GH",], horlines=FALSE, sizebyn=TRUE,
                pchcol=cols[8], add=TRUE)
  par(xpd=NA)
  legend(-160, -100, vegs_labels[1:4], fill=cols[1:4], bty='n')
  legend(-40, -100, vegs_labels[5:8], fill=cols[5:8], bty='n')

```

**Figure 1** Global distribution of studies included in BAAD. Symbols are sized relative to the number of individual plants in each dataset.

@@

plant traits; biomass allocation; biomass partitioning; plant allometry; allometric equations; global carbon cycle

@@

Count_obs <- function(x)length(data[[x]][!is.na(data[[x]])])
N <- sapply(thesevars, Count_obs)

Count_studies <- function(x)length(unique(data[["studyName"]][!is.na(data[[x]])]))
Nstud <- sapply(thesevars, Count_studies)

df <- rbind.fill(apply(data[,thesevars], 2, function(x)data.frame(Min=min(x, na.rm=T),Max=max(x, na.rm=T),Median=median(x, na.rm=T))))

dfr <- data.frame(Variable=Labels, N=N)
dfr <- cbind(dfr,df)

dfr <- dfr[dfr$N > 0,]
rownames(dfr) <- NULL

f <- function(x)formatC(x, digits=2, format="fg")
dfr$Min <- f(dfr$Min)
dfr$Max <- f(dfr$Max)
dfr$Median <- f(dfr$Median)

```

```{r results='asis', echo=FALSE}
options(digits=9)
kable(dfr, "markdown", align="l", padding=10)
```
